---
title: "Reproducible Research Activity Monitoring Project"
author: "Eric J Sullivan"
date: "June 5, 2019"
output: html_document
---

###Activity Monitoring Project for Reproducible Research Week 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(lattice)
```

1 <b>Code for reading in the dataset and/or processing the data</b>

```{r}
filename <- "ActivityMonitoring.zip"

#Download Files
if (!file.exists(filename)){
  fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(fileURL, filename, method="curl")
}  

#Unzip Files
if (!file.exists("activity.csv")) { 
  unzip(filename) 
}

ActivityData <- read.csv("activity.csv")
```

2 <b>Histogram of the total number of steps taken each day</b>

```{r}
DailySteps<-with(ActivityData, tapply(steps, date, FUN = sum))
hist(DailySteps,breaks = 10, main = "Histogram of Daily Steps", xlab = "Daily Steps")
```

3 <b>Mean and median number of steps taken each day

```{r}
MeanSteps <- mean(DailySteps, na.rm = TRUE)
MedianSteps <- median(DailySteps, na.rm = TRUE)
print(paste("Mean Steps =",MeanSteps))
print(paste("Median Steps =",MedianSteps))
```

4 <b>Time series plot of the average number of steps taken</b>

```{r}
StepsbyInterval<-with(ActivityData, tapply(steps, interval, FUN = mean, na.rm = TRUE))
plot(as.numeric(names(StepsbyInterval)),StepsbyInterval, main = "Average Steps by Time Interval", xlab = "Time Interval", ylab = "Steps", type = "l")
```

5 <b>The 5-minute interval that, on average, contains the maximum number of steps</b>

```{r}
MaxInterval <- StepsbyInterval[max(StepsbyInterval,na.rm = TRUE) == StepsbyInterval]
print(paste("The maximum average number of steps,", MaxInterval, ", occurs during interval", names(MaxInterval)))
```

6 <b>Code to describe and show a strategy for imputing missing data</b>

Here missing data will be replaced by the mean steps for that interval.

```{r}
print(paste("The total number of missing data points is", sum(is.na(ActivityData$steps))))
ImputedActivity <- ActivityData
ImputedActivity <- ImputedActivity %>% group_by(interval) %>% mutate(steps = ifelse(is.na(steps), round(mean(steps, na.rm = TRUE)), steps))
```


7 <b>Histogram of the total number of steps taken each day after missing values are imputed</b>

```{r}
ImputedDailySteps <- with(ImputedActivity, tapply(steps, date, FUN = sum))
hist(ImputedDailySteps, breaks = 10, main = "Histogram of Daily Steps with Imputed Values", xlab = "Daily Steps")
```
```{r}
ImputedMeanSteps <- mean(ImputedDailySteps, na.rm = TRUE)
ImputedMedianSteps <- median(ImputedDailySteps, na.rm = TRUE)
print(paste("Mean Steps with Imputed Values =",ImputedMeanSteps))
print(paste("Median Steps with Imputed Values =",ImputedMedianSteps))
```

8 <b>Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends</b>

```{r}
ImputedActivity <- mutate(ImputedActivity, DayofWeek = weekdays(as.Date(date, format = "%Y-%m-%d")), WeekPart = ifelse(DayofWeek %in%c("Saturday","Sunday"), "Weekend", "Weekday"))
WeekendImputedStepsbyInterval <- aggregate(steps~interval+WeekPart, data = ImputedActivity, FUN = "mean")
xyplot(steps~interval|WeekPart,WeekendImputedStepsbyInterval, layout = c(1,2), type = "l", xlab = "Interval", ylab = "Steps", main = "Average Steps per Interval for Each Part of the Week")
```

9 <b>All of the R code needed to reproduce the results (numbers, plots, etc.) in the report</b>